---
title: "p8105_hw6_rw3031"
author: "Polly Wu (rw3031)"
date: "2024-11-20"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(ggplot2)
library(httr)
library(modelr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

# Problem 1 

preparing the weather_dataframe 

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

bootstrapping the data 

```{r}
bootstrap =
weather_df|>
  modelr::bootstrap(n=5000)|>
  mutate(
    models = map(.x = strap , ~lm(tmax~tmin, data = .x))
  )
```

obtain the results from the model fit 

```{r}
bootstrap_result = 
bootstrap|>
  mutate(
    r_square = map(models, broom::glance),
    log = map(models, broom::tidy)
  )
```

## plot the distribution for r-squared 

```{r}
bootstrap_result|>
  unnest(r_square)|>
  select(r.squared)|>
  ggplot(aes(x=r.squared))+geom_histogram()
```
## the 95% confidence interval for r-squared 

```{r}
bootstrap_result|>
  unnest(r_square)|>
  select(r.squared)|>
  summarise(quantile_2.5 = quantile(r.squared, probs = 0.025),
            quantile_97.5 = quantile(r.squared, probs = 0.975))
```
The 95% confidence interval for r-squared is 0.8934 - 0.9275

```{r}
bootstrap_log = 
bootstrap_result|>
  unnest(log)|>
  select(.id, term, estimate)|>
  pivot_wider(
    names_from = term,
    names_prefix = "beta_",
    values_from = estimate
  )|>
  rename(beta_0 = "beta_(Intercept)",
         beta_1 = beta_tmin)|>
  mutate(log = log(beta_0*beta_1))
```

## plot for distribution of the log(beta_1*beta_0)

```{r}
bootstrap_log|>
  ggplot(aes(x=log))+geom_histogram()
```
## 95% confidence interval for log(beta_0*beta_1)

```{r}
bootstrap_log|>
  summarise(quantile_2.5 = quantile(log, probs = 0.025),
            quantile_97.5 = quantile(log, probs = 0.975))
```
The 95% confidence interval for log(beta_0*beta_1) is 1.9654 - 2.0581

# Problem 2

```{r}
homocide = 
  read_csv("./homicide-data.csv")|>
  janitor::clean_names()|>
  mutate(city_state = str_c(city, state, sep = ", "))|>
  filter(!city_state %in% c("Dallas, TX","Phoenix, AZ","Kansas City, MO", "Tulsa, AL"),
         victim_race %in% c("White", "Black"))|>
  mutate(victim_age = as.numeric(victim_age))
```
## fit glm model for baltimore 

```{r}
homocide|>
  filter(city_state == "Baltimore, MD")|>
  mutate(
    resolve = case_when(
      disposition == "Open/No arrest" ~ 0,
      TRUE ~ 1))|>
  summarise(model = list(glm(resolve~victim_age+victim_race+victim_sex)))|>
  mutate(results = map(model, ~ broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE)))|>
  unnest(results)|>
  select(term, estimate, conf.low, conf.high)|>
  filter(term == "victim_sexMale")
```

```{r}
homocide_glm = 
homocide|>
  mutate(
    resolve = case_when(
      disposition == "Open/No arrest" ~ 0,
      TRUE ~ 1))|>
  group_by(city_state)|>
  nest()|>
  mutate(model = map(.x=data, ~glm(resolve~victim_age+victim_race+victim_sex, 
                                         data = .x)),
         results = map(model, ~ broom::tidy(.x, conf.int = TRUE, exponentiate = TRUE)))|>
  unnest(results)|>
  select(term, estimate, conf.low, conf.high)|>
  filter(term == "victim_sexMale")|>
  select(-term)

homocide_glm
```

## plot for the or estimate and confidence interval on homocide for each city

```{r}
homocide_glm|>
  arrange(desc(estimate))|>
  ggplot(aes(y=fct_reorder(city_state,estimate), x=estimate))+
  geom_point() +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high))+
  labs(
    x = "City, State",
    y = "Odds ratio estimate and 95% confidence interval"
  )
```

# Problem 3

## import the birthweight data 

```{r}
birthweight = 
  read_csv("./birthweight.csv")

summary(birthweight)
```

## clean the birthweight data for analysis

I convert the categorical variables from numeric to factor and remove the variables with complete missing values 

```{r}
birthweight =
birthweight|>
  mutate(babysex = factor(babysex),
         frace = factor(frace),
         malform = factor(malform),
         mrace = factor(mrace))|>
  select(-pnumlbw,-pnumsga)


summary(birthweight)
```
## build my regression model for birthweight

My assumption is that child birthweight is going to be associated with gestational age in weeks, mother's race, and average number of cigarettes smoked per day during pregnancy. 

```{r}
fit_bw = lm(bwt ~ gaweeks+mrace+smoken, data=birthweight)

fit_bw|>
  broom::tidy()|>
  knitr::kable(digits = 4)
```



## getting the residual plot

```{r}
birthweight|>
  modelr::add_residuals(fit_bw)|>
  modelr::add_predictions(fit_bw)|>
  ggplot(aes(x=pred, y= resid))+geom_point(alpha = 0.3)
```

Since the residual is quite randomly scattered around 0, we conclude the overall fit of the model is good. 

## build the model using birth length and gestational age

```{r}
fit_length = lm(bwt ~ blength+gaweeks, data=birthweight)

fit_length|>
  broom::tidy()|>
  knitr::kable(digits = 4)
```

## build the model for head circumference, length, and sex

```{r}
fit_headlengthsex = lm(bwt ~ bhead+blength+babysex
                       +bhead*blength+blength*babysex+babysex*bhead+bhead*blength*babysex,
                       data=birthweight)

fit_headlengthsex|>
  broom::tidy()|>
  knitr::kable(digits = 4)
```

## cross validation

```{r}
cv_df =
  crossv_mc(birthweight, 100)
```


```{r}
cv_df =
cv_df|>
  mutate(model_gaweeks = map(train, \(df) lm(bwt ~ gaweeks+mrace+smoken, data=df)),
         model_length = map(train, \(df) lm(bwt ~ blength+gaweeks, data=df)),
         model_headlengthsex = map(train, \(df) lm(bwt ~ bhead+blength+babysex
                       +bhead*blength+blength*babysex+babysex*bhead+bhead*blength*babysex,
                       data=df))
           )|>
  mutate(
    rmse_gaweeks = map2_dbl(model_gaweeks, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_length = map2_dbl(model_length, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_headlengthsex = map2_dbl(model_headlengthsex, test, \(mod, df) rmse(model = mod,      data = df))
  )
```

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

Based on the rootmean square models, the headlengthsex model has the smallest rmse thus we conclude that to be the best model of the three models. 
